---
title: "Math 4753 Lab 1 Introduction to R"
author: "Vincent Tran"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_caption: yes
    fig_height: 4
    fig_width: 5
    number_sections: yes
    toc: true
    toc_float: true
    toc_collapsed: false
  pdf_document: default
---
# Lab 1

## Task 1: get working directory

```{r}
getwd()
```

## Task 2: save excel as csv

```{r}
ddt = read.csv("ddt.csv")
head(ddt)
```
## Task 3: Subset data
```{r}
summary(ddt)
```

Qualitative Variables:
- RIVER
- SPECIES
Quantitative Variables:
- MILE
- LENGTH
- WEIGHT
- DDT

Number of species in dataset: 144

Large mouth basses with weight bigger than 800gms
```{r}
ddtHeavyLmbass = ddt[ddt$SPECIES == "LMBASS" & ddt$WEIGHT > 800, ]
ddtHeavyLmbass
```

SCM river where DDT bigger than 4.0
```{r}
ddtScmRiverDdt4 = ddt[ddt$RIVER == "SCM" & ddt$DDT > 4.0, ]
ddtScmRiverDdt4
```

## Task 4: 4 questions
1. mean length of fish in ddt: 42.8125
```{r}
meanFishLengthDdt = mean(ddt$LENGTH)
meanFishLengthDdt
```
2. standard deviation of fish weight in ddt: 376.5461
```{r}
sd(ddt$WEIGHT)
```
Q3. Is this the plot of LENGTH vs WEIGHT from the ddt data frame? No | 0
It is a WEIGHT vs LENGTH plot.
```{r}
library(ggplot2)

plot <- ggplot(ddt, aes(x=WEIGHT, y=LENGTH)) + geom_point()
plot
```

Q4. If v=1:20 what is the last value of v/20?: 1
```{r}
v=1:20
v/20
```



## Task 5: Tables and bar plots

River table and bar plot
```{r}
riverTable <- with(ddt,table(RIVER))
riverTable
barplot(riverTable, beside=TRUE, col=1:4)
```

River species table and bar plot
```{r}
riverSpeciesTable=with(ddt,table(RIVER,SPECIES))
riverSpeciesTable

barplot(riverSpeciesTable, beside=TRUE, col=rainbow(4))
```

## Task 6: Pie charts
species pie chart
```{r}
speciesTable <- with(ddt,table(SPECIES))
pie(speciesTable,col=1:3)
```

rivers pie chart
```{r}
riverTable <- with(ddt,table(RIVER))
pie(riverTable, col = 1:4)
```

## Task 7: boxplot

DDT boxplot
```{r}
ggplot(ddt, aes(DDT)) + geom_boxplot()
```

WEIGHT boxplot
```{r}
ggplot(ddt, aes(WEIGHT)) + geom_boxplot()
```

LENGTH boxplot
```{r}
ggplot(ddt, aes(LENGTH)) + geom_boxplot()
```

## Task 8: coplot
LENGTH Vs WEIGHT given RIVER coplot
```{r}
rivcol=with(ddt, ifelse(RIVER=="FCM","Red",
                        ifelse(RIVER=="LCM","Blue",
                               ifelse(RIVER=="SCM","Green","Black"))))
coplot(LENGTH~WEIGHT|RIVER,data=ddt, col = rivcol)
```

DDT Vs WEIGHT given SPECIES coplot
```{r}
fishcol=with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))
coplot(DDT~WEIGHT|SPECIES, data=ddt, col=fishcol)
```


## Task 9: Replicating plots with ggplot

Plot 1
vertical boxplot weight vs species by river with legend
```{r}
VincentTran <- ggtitle("Vincent Tran")
barchartPlotWeightSpeciesRiver <- ggplot(ddt, aes(x=SPECIES, y=WEIGHT, fill=RIVER)) + geom_boxplot()

barchartPlotWeightSpeciesRiver + VincentTran
```

Plot 2
```{r}
violinPlotLengthRiverSpecies <- ggplot(ddt, aes(x=RIVER, y=LENGTH, fill=SPECIES)) + geom_violin()

violinPlotLengthRiverSpecies + VincentTran
```


Plot 3
```{r}
scatterPlotLengthWeightSpecies <- ggplot(ddt, aes(x=WEIGHT, y=LENGTH, color=SPECIES)) + geom_point()

scatterPlotLengthWeightSpecies + VincentTran
```

